Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-03-20 paper.tex
Transition to ASC Nano complete

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Robert Jensen}
\author{Thomas Andersen}
\author{Anders Nierhoff}
\altaffiliation{CINF, Department of Physics, Technical University of Denmark, Fysikvej 312, 2800 Lyngby, Denmark.}
\author{Thomas Pedersen}
\author{Ole Hansen}
\affiliation{CINF, Department of Micro- and Nanotechnology, Technical University of Denmark, Ørsteds Plads, B 345B, 2800 Lyngby, Denmark. Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: ibchork@fysik.dtu.dk}
\author{Søren Dahl}
\author{Ib Chorkendorff}
\email{ibchork@fysik.dtu.dk}
\altaffiliation{CINF, Department of Physics, Technical University of Denmark, Fysikvej 312, 2800 Lyngby, Denmark.}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{reactor.jpg}
  \caption{Image of the 20x16\,mm microreactor showing the circular reactor area and gas channel system}
  \label{fgr:reactor}
\end{figure}

@@

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup directly after cluster deposition and anodic bonding. Initially a light-off CO-oxidation measurement is performed (shown in Figure~\ref{fgr:initial_treatment}) where the temperature is increased at 3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to CO$_2$. At CO:O$_2$ ratio of $\sim0.08$ and a geometrical Pt coverage of $\sim0.1\%$ light-off occurs at approximately 180\,$^\circ$C. The temperature ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is decreased at the same rate as the heating ramp. After the initial light-off the sample is cooled to room temperature. When at room temperature the temperature is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical light-off ramp and the subsequent increase in temperature to achieve sustained oscillations is seen. Often the sample will show oscillations either directly upon light-off or on the falling temperature ramp.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After performing a light-off ramp where oscillations can been seen the temperature is increased to a constant value where sustained oscillations take place. The non-zero value of CO during high conversion periods is consistent with the expected QMS background signals from CO$_2$ and O$_2$.}
  \label{fgr:initial_treatment}
\end{figure}
Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm particles with a geometrical coverage of 0.1\% of the reactor area gave the most stable oscillations. No consistent dependency on oscillation frequency and duty cycle was found that could be attributed to the size of the particles.

@@

An example of a single very slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where all CO is converted to CO$_2$ (notice the CO signal does not go to zero due to the cracking pattern of CO$_2$ and CO background in the QMS) the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases slowly over time until approximately 50\% conversion is reached (at $\sim$1118\,min). Shortly after reaching 50\% conversion the sample again ignites and convert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate deactivation. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of careful parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,min. Inserted circles illustrates the proposed model. At high conversion the platinum is oxidized (red) while in low conversion the sample is reduced (blue).}
  \label{fgr:full_oscillation}
\end{figure*}

After prolonged measurements over several days an increase in oscillation period was observed as shown in Figure \ref{fgr:long_measurement}. Initially, the oscillation period is $\sim$2.5 min and increases linearly with time until 1500 min of total oscillation time. Hereafter the oscillations become more irregular as shown in Figure~\ref{fgr:extracts} and lower panel Figure~\ref{fgr:long_measurement}. At experiment end the oscillation period is between 20 and 45 min.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample oscillating under constant temperature, pressure and reactant composition. The sample went through a total of 439 oscillations in 4 days. Time is defined from experiment start and hence includes initial treatment. The inset shows the initial steady increase in oscillation period.}
  \label{fgr:long_measurement}
\end{figure}

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations extracted from the same four days measurement. Data from shortly after the initial treatment, after 2800\,min of oscillation time and after 5900\,min are shown. After 5900\,min the oscillations are more irregular which is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also, oscillations in between full conversion oscillations with smaller amplitude and much higher frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 days long experiment. The oscillation period becomes more irregular with time while the total integrated conversion remains constant. Furthermore, small oscillations in between full conversion cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{figure}

@@

%Furthermore, the oscillations have the qualitatively identical behavior as shown in Figure~\ref{fgr:full_oscillation}. A short full conversion period is followed by a fast deactivation from where reactivity is slowly recovered. At approximately 50\% conversion light-off occurs and the sample is in full conversion.

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{temperature_dependence.png}
  \caption{After about 10 hours of oscillations at 205$^\circ$C the temperature is first lowered by 5$^\circ$C and soon after increased by 2$^\circ$C. Before the temperature step the oscillation period was approximately 500\,s and after the step the period is approximately 5000\,s.}
  \label{fgr:temperature_dependence}
\end{figure}
A very strong temperature dependence on the oscillation frequency was also found as shown in Figure~\ref{fgr:temperature_dependence}. By changing the temperature 5$^\circ$C the oscillation period was changed by more than a factor of 10. The magnitude of the change is not completely consistent across all measured samples but all samples show a very large temperature dependency where in some cases a provoked small change in temperature will consistently turn on and off the oscillations.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{oscillations_gas_dependence_summary_supplemental.png}
  \caption{Summary of the gas-dependence. The overall trend of increasing oscillation period is now supporimposed by small discontinous stpes when the CO concentration is increased.}
  \label{fgr:gas_dependence_summary}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[width=14cm]{oscillations_gas_dependence_supplemental.png}
  \caption{Complete series of oscillating sample, showing CO (red) and CO$_2$ (blue). The temperature is constant $210^\circ$C doing the entire measurement. For every frame the CO-concentration is increased.}
  \label{fgr:gas_dependence}
\end{figure*}

@@

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{temperature_dependence_supplemental.png}
  \caption{Another example of the very pronounced temperature dependence.}
  \label{fgr:temperature_dependence_supplemental}
\end{figure}

@@

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{duty_cycles_long_measurement_supplemental.png}
  \caption{A plot of the duty-cycle of the sample duing the 4-day long experiment. The individual data-points is calculted as the integral of CO (red) or CO$_2$ (blue) doing the individual oscillations. The ratio between CO and CO$_2$ is drawn in black.}
  \label{fgr:duty_cycles_supplemental}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Robert Jensen}
\author{Thomas Andersen}
\author{Anders Nierhoff}
\affiliation{CINF, Department of Physics, Technical University of Denmark, Fysikvej 312, 2800 Lyngby, Denmark.}
\author{Thomas Pedersen}
\author{Ole Hansen}
\affiliation{CINF, Department of Micro- and Nanotechnology, Technical University of Denmark, \O rsteds Plads, B 345B, 2800 Lyngby, Denmark.}
\author{S\o ren Dahl}
\author{Ib Chorkendorff}
\email{ibchork@fysik.dtu.dk}
\affiliation{CINF, Department of Physics, Technical University of Denmark, Fysikvej 312, 2800 Lyngby, Denmark.}

@@

\begin{scheme}
  \includegraphics[width=8cm]{reactor.jpg}
  \caption{Image of the 20x16\,mm microreactor showing the circular reactor area and gas channel system}
  \label{fgr:reactor}
\end{scheme}

@@

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup directly after cluster deposition and anodic bonding. Initially a light-off CO-oxidation measurement is performed (shown in Figure~\ref{fgr:initial_treatment}) where the temperature is increased at 3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to CO$_2$. At CO:O$_2$ ratio of $\sim0.08$ and a geometrical Pt coverage of $\sim0.1\%$ light-off occurs at approximately 180\,$^\circ$C. The temperature ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is decreased at the same rate as the heating ramp. After the initial light-off the sample is cooled to room temperature. When at room temperature the temperature is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical light-off ramp and the subsequent increase in temperature to achieve sustained oscillations is seen. Often the sample will show oscillations either directly upon light-off or on the falling temperature ramp.

\begin{scheme}
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After performing a light-off ramp where oscillations can been seen the temperature is increased to a constant value where sustained oscillations take place. The non-zero value of CO during high conversion periods is consistent with the expected QMS background signals from CO$_2$ and O$_2$.}
  \label{fgr:initial_treatment}
\end{scheme}

Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm particles with a geometrical coverage of 0.1\% of the reactor area gave the most stable oscillations. No consistent dependency on oscillation frequency and duty cycle was found that could be attributed to the size of the particles.

@@

An example of a single very slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where all CO is converted to CO$_2$ (notice the CO signal does not go to zero due to the cracking pattern of CO$_2$ and CO background in the QMS) the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases slowly over time until approximately 50\% conversion is reached (at $\sim$1118\,min). Shortly after reaching 50\% conversion the sample again ignites and convert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself.

\begin{scheme}
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate deactivation. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of careful parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,min. Inserted circles illustrates the proposed model. At high conversion the platinum is oxidized (red) while in low conversion the sample is reduced (blue).}
  \label{fgr:full_oscillation}
\end{scheme}

After prolonged measurements over several days an increase in oscillation period was observed as shown in Figure \ref{fgr:long_measurement}. Initially, the oscillation period is $\sim$2.5 min and increases linearly with time until 1500 min of total oscillation time. Hereafter the oscillations become more irregular as shown in Figure~\ref{fgr:extracts} and lower panel Figure~\ref{fgr:long_measurement}. At experiment end the oscillation period is between 20 and 45 min.

\begin{scheme}
  \includegraphics[width=9cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample oscillating under constant temperature, pressure and reactant composition. The sample went through a total of 439 oscillations in 4 days. Time is defined from experiment start and hence includes initial treatment. The inset shows the initial steady increase in oscillation period.}
  \label{fgr:long_measurement}
\end{scheme}

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations extracted from the same four days measurement. Data from shortly after the initial treatment, after 2800\,min of oscillation time and after 5900\,min are shown. After 5900\,min the oscillations are more irregular which is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also, oscillations in between full conversion oscillations with smaller amplitude and much higher frequency than the full on-off cycles become more visible as time progresses.
\begin{scheme}
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 days long experiment. The oscillation period becomes more irregular with time while the total integrated conversion remains constant. Furthermore, small oscillations in between full conversion cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{scheme}

@@

\begin{scheme}
  \includegraphics[width=9cm]{temperature_dependence.png}
  \caption{After about 10 hours of oscillations at 205$^\circ$C the temperature is first lowered by 5$^\circ$C and soon after increased by 2$^\circ$C. Before the temperature step the oscillation period was approximately 500\,s and after the step the period is approximately 5000\,s.}
  \label{fgr:temperature_dependence}
\end{scheme}

A very strong temperature dependence on the oscillation frequency was also found as shown in Figure~\ref{fgr:temperature_dependence}. By changing the temperature 5$^\circ$C the oscillation period was changed by more than a factor of 10. The magnitude of the change is not completely consistent across all measured samples but all samples show a very large temperature dependency where in some cases a provoked small change in temperature will consistently turn on and off the oscillations.

@@

\begin{scheme}
  \includegraphics[width=9cm]{oscillations_gas_dependence_summary_supplemental.png}
  \caption{Summary of the gas-dependence. The overall trend of increasing oscillation period is now supporimposed by small discontinous stpes when the CO concentration is increased.}
  \label{fgr:gas_dependence_summary}
\end{scheme}

\begin{scheme}
  \includegraphics[width=14cm]{oscillations_gas_dependence_supplemental.png}
  \caption{Complete series of oscillating sample, showing CO (red) and CO$_2$ (blue). The temperature is constant $210^\circ$C doing the entire measurement. For every frame the CO-concentration is increased.}
  \label{fgr:gas_dependence}
\end{scheme}

@@

\begin{scheme}
\centering
  \includegraphics[width=9cm]{temperature_dependence_supplemental.png}
  \caption{Another example of the very pronounced temperature dependence.}
  \label{fgr:temperature_dependence_supplemental}
\end{scheme}

@@

\begin{scheme}
  \includegraphics[width=9cm]{duty_cycles_long_measurement_supplemental.png}
  \caption{A plot of the duty-cycle of the sample duing the 4-day long experiment. The individual data-points is calculted as the integral of CO (red) or CO$_2$ (blue) doing the individual oscillations. The ratio between CO and CO$_2$ is drawn in black.}
  \label{fgr:duty_cycles_supplemental}
\end{scheme}
